---
title: "PCT-guided stewardship workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PCT-guided stewardship workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

> **Research and educational use only.** Not a medical device. All data in this
> vignette are synthetic.

```{r setup}
library(pctsteward)
```

This vignette walks through the package end to end: classifying a single PCT
value, following a patient's PCT series to a stop decision, screening a renal
dose, and summarising a simulated cohort.

## 1. Classifying a PCT value

The decision bands come from one constant, `pct_cutoffs`. A value exactly on a
cut-off belongs to the higher band.

```{r}
pct_cutoffs
classify_pct(c(0.05, 0.1, 0.3, 0.5, 4))
```

## 2. Following PCT over time

`pct_clearance()` gives the running decline from the highest value so far.
`asp_recommendation()` turns each timepoint into Continue, De-escalate or
Discontinue.

```{r}
responder <- c(2.5, 1.8, 0.4, 0.15, 0.08)
round(pct_clearance(responder), 2)
asp_recommendation(responder)

non_responder <- c(2.5, 2.7, 2.4, 2.6, 2.3)
asp_recommendation(non_responder)
```

`pct_decline_from_peak()` summarises the latest state of a series:

```{r}
str(pct_decline_from_peak(responder))
```

By default the stop signal also fires when PCT falls below 0.25 ng/mL, even
without an 80% decline. Setting `abs_stop = 0` uses the 80% rule alone:

```{r}
pct_decline_from_peak(c(0.3, 0.2))$stop_signal
pct_decline_from_peak(c(0.3, 0.2), abs_stop = 0)$stop_signal
```

The kinetics plot colours each patient by whether the rule is met:

```{r kinetics, fig.alt = "PCT curves for two patients over 7 days: one falls below 0.1 ng/mL and meets the stop rule; the other stays near 2.5 ng/mL and does not."}
plot_pct_kinetics()
```

## 3. Renal dose screening

`screen_medications()` estimates creatinine clearance with Cockcroft-Gault,
choosing actual, ideal or adjusted body weight, and compares the ordered
regimen with `renal_dose_rules`.

```{r}
renal_dose_rules

res <- screen_medications(
  age_yrs = 68, weight_kg = 105, height_cm = 160, scr_mgdl = 1.8,
  sex_is_female = TRUE, drug_name = "cefepime", prescribed_regimen_id = "2g_q8h"
)
res[c("weight_model", "dosing_weight_kg", "crcl_ml_min", "recommended_label", "status")]
```

The helpers are vectorised, so a whole ward can be screened at once:

```{r}
ward <- data.frame(
  age = c(45, 72, 60), weight = c(70, 95, 55), height = c(172, 165, 158),
  scr = c(0.9, 2.1, 3.5), female = c(FALSE, TRUE, TRUE)
)
ibw <- ideal_body_weight(ward$height, ward$female)
ward$crcl <- round(cockcroft_gault(ward$age, dosing_weight(ward$weight, ibw),
                                   ward$scr, ward$female), 1)
ward$regimen <- recommend_regimen("cefepime", ward$crcl)
ward
```

## 4. Cohort view

`simulate_cohort()` builds a synthetic long-format dataset matching
`DATA_DICTIONARY.md`: one row per patient per `Time_Hour`.

```{r}
cohort <- simulate_cohort(n = 1000, seed = 2026)
head(cohort[, c("Patient_ID", "Time_Hour", "PCT_Level", "PCT_Clearance",
                "ASP_Recommendation")], 8)
```

How recommendations shift over the first 72 hours:

```{r}
round(prop.table(table(cohort$Time_Hour, cohort$ASP_Recommendation), 1), 2)
```

The heatmap uses each patient's admission row only, and hides mortality for
cells with fewer than `min_cell_n` patients:

```{r heatmap, fig.height = 5, fig.alt = "Heatmap of 30-day mortality by admission PCT band and baseline SOFA band; mortality rises towards high PCT and high SOFA."}
build_stewardship_heatmap(cohort)
```
